Batched Bayesian Optimization via Multi-objective Acquisition Ensemble for Automated Analog Circuit Design

### 2018-02-05 paper/background.tex
Paper update

We handle the scenarios where topology of the analog circuit is fixed, this is
practical as thare are usually a lot of classical topologies for given design
task. Once the circuit topology is fixed, the designer has to choose the
appropriate design parameters according to the specifications and the circuit
device model. What we want to do is to automatically search for the optimal
design parameters, the problem can then be formulated as a bound-constrained
black-box optimization problem:

@@

where $\bm{x} \in \textrm{D} \subset \textrm{R}^d$ is the design variables
and $f_i(\bm{x})$ represents the $i$-th cared circuit performance like gain
or phase margin of an amplifier. and the $\mathrm{FOM}(\bm{x})$ is the
design objective. Given the design parameters $\bm{x}$, the FOM value can be
obtained via circuit simulation using softwares like HSPICE or spectre.

% TODO: Explain the noise

@@

We model the $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} with Gaussian
process (GP) model~\cite{GPML}. The GP model is the most commonly used model
for Bayesian optimization, the advantage of GP is that it can give a
well-calibrated uncertainty of prediction. GP is characterized by a mean
function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In this
paper, we use squared-exponential ARD kernel, and a constant mean function
$m(\bm{x}) = \mu_0$ for all our experiments.

@@

Bayesian optimization~\cite{shahriari2016taking} was proposed for the
optimization of expensive black-box functions. It consists of two essential
ingredients, i.e., the probabilistic surrogate models and an acquisition
function. The probabilistic surrogate models provide the prediction with
uncertainties. They are refined incrementally with newly observed data.
Acquisition function is used to explore the state space based on the surrogate
model optimally. The procedure of Bayesian optimization is described in
Algorithm~\ref{alg:BayesianOptAlgo}.

\begin{algorithm}
\caption{Bayesian Optimization}
\label{alg:BayesianOptAlgo}
\begin{algorithmic}[1]
\STATE Initial Sampling
\STATE Construct initial GP model
\FOR{t = 1, 2, \dots}
\STATE Construct the acquisition function
\STATE Find $\bm{x}_t$ that optimizes the acquisition function
\STATE Sample $y_t = f(\bm{x}_t)$
\STATE Update probabilistic surrogate model
\ENDFOR
\STATE Return best $f(\bm{x})$ recorded during iterations
\end{algorithmic}
\end{algorithm}

@@

Following the suggestion of~\cite{brochu2010tutorial}, the $\kappa$ in \eqref{eq:LCB} is defined as:
\begin{equation}
    \label{eq:LCB}
    \begin{array}{lll}
        \kappa &=& \sqrt{\nu \tau_t} \\
        \tau_t &=& 2 \log(t^{d/2+2} \pi^2 / 3 \delta)
    \end{array}
\end{equation}
Where $t$ is the number of current iteration. $\nu$ and $\delta$ are two user defined parameters, we fix $\nu = 0.5$ and $\delta = 0.05$ in this paper for the proposed MACE algorithm and our implementation of the BLCB algorithm.

@@

There are also other acquisition function, like the knowledge gradient~\cite{scott2011correlated} function, predictive entropy search~\cite{hernandez2014predictive}. A portforlio of several acquisition function is also possible~\cite{hoffman2011portfolio}.

>>>>

We handle the scenarios where the topology of the analog circuit is fixed, this
is practical as there are usually a lot of classical topologies for given
design task. Once the circuit topology is fixed, the designer has to choose the
appropriate design parameters according to the specifications and the circuit
device model. What we want to do is automatically searching for the optimal
design parameters, the problem can then be formulated as a bound-constrained
black-box optimization problem:

@@

where $\bm{x} \in \textrm{D}$ is the vector of design variables,
$\mathrm{FOM}(\bm{x})$ is the objective constructed from the design
specifications. Given the design parameters $\bm{x}$, the FOM value can be
obtained by commercial circuit simulators like HSPICE or spectre.

\textcolor{red}{TODO: If there is still space, illustrate the objecive with the PM-width figure}

\begin{figure}[]
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/PM.eps}}
        \caption{Phase margin of an amplifier as function of the width of a transistor}
        \label{fig:PMOpAmp}
    \end{center}
\end{figure}

@@

\textcolor{red}{
    TODO:
    \begin{enumerate}
            \item Noise
            \item MLE learning
    \end{enumerate}
}

We model the $\mathrm{FOM}(\bm{x})$ in \eqref{eq:Formulation} with Gaussian
process (GP) model~\cite{GPML}. The GP model is the most commonly used model
for Bayesian optimization, the advantage of GP is that it can give a
well-calibrated uncertainty of prediction. GP is characterized by a mean
function $m(\bm{x})$ and a covariance function $k(\bm{x}, \bm{x'})$. In this
work, we use squared-exponential ARD kernel, and a constant mean function
$m(\bm{x}) = \mu_0$ for all our experiments.

@@

Bayesian optimization~\cite{shahriari2016taking} was proposed for the
optimization of expensive black-box functions, it has two essential
ingredients, i.e., the probabilistic surrogate models and acquisition
functions. The probabilistic surrogate models provide predictions with
uncertainties. The acquisition functions make use of the predictive
distribution to explore the state space, the procedure of Bayesian optimization
is summarized in Algorithm~\ref{alg:BayesianOptAlgo}.

\textcolor{red}{TODO: Add Input, Output parameters}
\begin{algorithm}
\caption{Bayesian Optimization}
\label{alg:BayesianOptAlgo}
\begin{algorithmic}[1]
\STATE Initial Sampling
\STATE Construct initial GP model
\FOR{t = 1, 2, \dots}
\STATE Construct the acquisition function
\STATE Find $\bm{x}_t$ that optimizes the acquisition function
\STATE Sample $y_t = f(\bm{x}_t)$
\STATE Update probabilistic surrogate model
\ENDFOR
\STATE Return best $f(\bm{x})$ recorded during iterations
\end{algorithmic}
\end{algorithm}

@@

Following the suggestion of~\cite{brochu2010tutorial}, the $\kappa$ in \eqref{eq:LCB} is defined as:
\begin{equation}
    \label{eq:LCBKappa}
    \begin{array}{lll}
        \kappa &=& \sqrt{\nu \tau_t} \\
        \tau_t &=& 2 \log(t^{d/2+2} \pi^2 / 3 \delta)
    \end{array}
\end{equation}
where $t$ is the number of current iteration. $\nu$ and $\delta$ are two user-defined parameters, we fix $\nu = 0.5$ and $\delta = 0.05$ in this paper for the proposed MACE algorithm and our implementation of the BLCB algorithm.

@@

There are also other acquisition functions, like the knowledge gradient~\cite{scott2011correlated} function, predictive entropy search~\cite{hernandez2014predictive}. A portfolio of several acquisition functions is also possible~\cite{hoffman2011portfolio}.

### 2018-02-05 paper/experiments.tex
Paper update

@@

We tested the MACE algorithm and other parallel BO method using eight commonly used benchmark
functions, as summarized in Table~ref{tab:summaryanalygical}.

@@

For all functions except the two 10D functions, we set the number of initial
random sampling to $N_{init} = 20$ and the number of iterations to $N_{iter} =
45$. Batch size are set to $B = 4$, the total number of function evaluations is
$N_{init} + B \times N_{iter}$. For the 10D ackley and 10D rosenbrock functions, we
$N_{init} = 100$ and $N_{iter} = 175$. The experiments were repeated ten
times to average the random fluctuations.

We also compared MACE algorithm with the EI and LCB acquisition functions in
sequential mode, the sequential EI and LCB are implemented by setting the batch
size $B = 1$ for EI-LP and BLCB algorithms.

The mean convergence plot are shown in Figure~\ref{fig:CovPlotBenchmark}.

As can be seen in Figure~\ref{fig:CovPlotBenchmark}, when running with batch
size $B = 4$, the MACE algorithm gives the best performance for six out of the
eight benchmarks. For the 2D Ackley function, EI-LP is the best algorithm,
while for the Eggholder function, the MACE, BLCB and EI-LP have quite similar
performance. Compare the batched MACE and the sequential MACE, the speedup is
also dramatic.

% \begin{table*}[htbp]
%     \centering
%     \caption{Optimization results of the benchmark functions\textcolor{red}{Alert: data still incomplete}}
%     \label{tab:result_analytical}
%     \begin{tabular}{lllllllllll}
%         \toprule
%         Function           & MACE-1             & LCB-1             & EI-1            & MACE-4             & BLCB-4  & EI-LP-4 & qKG-4 & qEI-4  \\ \midrule
%          Branin            & 0.3979$\pm$1.31e-5 & 0.3980$\pm$1.1e-4 & 0.414$\pm$0.016 & 0.3979$\pm$6.64e-6 &         &         &       &        \\
%          Alpine1           & 2.66$\pm$1.06      & 5.67$\pm$1.77     & 2.46$\pm$1.56   &                    &         &         &       &        \\
%          Hartmann6         & -3.26$\pm$0.06     & -3.20$\pm$0.12    & -3.21$\pm$0.15  &                    &         &         &       &        \\
%          Ackley2           & 1.72$\pm$1.12      & 1.62$\pm$0.926    & 1.01$\pm$0.98   &                    &         &         &       &        \\
%          Rosenbrock2       & 0.03$\pm$0.05      & 0.03$\pm$0.02     & 13.55$\pm$9.527 &                    &         &         &       &        \\
%          Ackley10          &                    &                   &                 &                    &         &         &       &        \\
%          Rosenbrock10      &                    &                   &                 &                    &         &         &       &        \\
%         \bottomrule
%     \end{tabular}
% \end{table*}

@@

\begin{figure}[htbp]
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/sopam.pdf}}
        \caption{Schematic of the operational amplifier}
        \label{fig:schDAC2014}
    \end{center}
\end{figure}

@@

For this circuit, we compared the MACE algorithm with the BLCB and EI-LP
algorithms, the qKG and qEI are not compared as the calculation of qEI and qKG
acquisition functions become very slow for the ten dimensionl functions.

@@

\begin{figure}[htbp]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{./img/mean_DAC2014.eps}}
\caption{Optimization results of the operational amplifier}
\label{fig:resDAC2014}
\end{center}
\vskip -0.2in
\end{figure}

@@

For this power amplifier, we aim to maximize the power added effiency (PAE) and the output power (Pout), the $FOM$ is constructed as
$$
\mathit{FOM} = -3 * \mathit{PAE} - \mathit{Pout}
$$

The MACE, BLCB and EI-LP algorithms were tested in both, sequential and batched
mode, the number of initial sampling is $N_{init} = 100$, the number of
iterations is $N_{iter} = 100$, the batch size is set to $B = 4$, so the total
number of HSPICE simulations is 500 for each batched run and 200 for each
sequential run.

@@

\begin{figure*}[]
    \begin{center}
        \centerline{\includegraphics[width=1.0\linewidth]{./img/convplot.eps}}
        \caption{Optimization results of the benchmark functions}
        \label{fig:CovPlotBenchmark}
    \end{center}
\end{figure*}

>>>>

\begin{figure*}[htbp]
    \begin{center}
        \centerline{\includegraphics[width=1.0\linewidth]{./img/convplot.eps}}
        \caption{Optimization results of the benchmark functions}
        \label{fig:CovPlotBenchmark}
    \end{center}
\end{figure*}

@@

We tested the MACE algorithm and other parallel BO method using eight commonly used benchmark
functions, as summarized in Table~\ref{tab:summaryanalygical}.

@@

For all functions except the two 10D functions, we set the number of initial
random sampling to $N_{init} = 20$ and the number of iterations to $N_{iter} =
45$. Batch size is set to $B = 4$, the total number of function evaluations is
$N_{init} + B \times N_{iter}$. For the 10D Ackley and 10D Rosenbrock functions, we
$N_{init} = 100$ and $N_{iter} = 175$. The experiments were repeated ten
times to average the random fluctuations.

We also ran the MACE algorithm in sequential mode and compared the EI and LCB
acquisition functions, the sequential EI and LCB based Bayesian optimization
are implemented by setting the batch size $B = 1$ for EI-LP and BLCB
respectively.

The mean convergence plot is shown in Figure~\ref{fig:CovPlotBenchmark}. The
statistics of the final optimization results are listed in
Table~\ref{tab:result_analytical}. As can be seen, when running with batch size
$B = 4$, the MACE algorithm gives the best performance for six out of the eight
benchmarks.  For the 2D Ackley function, EI-LP is the best algorithm, while for
the Eggholder function, the MACE, BLCB, and EI-LP have quite similar
performance.  Compare the batched MACE and the sequential MACE, the speedup is
also dramatic.

\begin{table*}[tp]
    \centering
    \caption{Optimization results of the benchmark functions}
    \label{tab:result_analytical}
    \begin{tabular}{lllllllllll}
        \toprule
        Function           & MACE-1 & LCB-1  & EI-1  & MACE-4  & BLCB-4  & EI-LP-4 & qKG-4 & qEI-4  \\ \midrule
         Branin            &        &        &       &         &         &         &       &        \\
         Alpine1           &        &        &       &         &         &         &       &        \\
         Hartmann6         &        &        &       &         &         &         &       &        \\
         Eggholder         &        &        &       &         &         &         &       &        \\
         Ackley2           &        &        &       &         &         &         &       &        \\
         Rosenbrock2       &        &        &       &         &         &         &       &        \\
         Ackley10          &        &        &       &         &         &         &       &        \\
         Rosenbrock10      &        &        &       &         &         &         &       &        \\
        \bottomrule
    \end{tabular}
\end{table*}

@@

\textcolor{red}{TODO: if possible, run qKG and qEI for this circuit}

@@

For this circuit, we compared the MACE algorithm with the BLCB and EI-LP
algorithms, the qKG and qEI are not compared as the calculation of qEI and qKG
acquisition functions become very slow for the ten-dimensional functions.

\begin{figure}[h]
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/sopam.pdf}}
        \caption{Schematic of the operational amplifier}
        \label{fig:schDAC2014}
    \end{center}
\end{figure}

\begin{figure}[]
\vskip 0.2in
\begin{center}
\centerline{\includegraphics[width=\columnwidth]{./img/mean_DAC2014.eps}}
\caption{Optimization results of the operational amplifier}
\label{fig:resDAC2014}
\end{center}
\vskip -0.2in
\end{figure}

@@

\begin{table}[]
    \centering
    \caption{Optimization results of the operational amplifier}
    \label{tab:result_opamp}
    \begin{tabular}{llll}
        \toprule
        Algorithm & Results     \\ \midrule
        MACE-1    & 0.0$\pm$0.0 \\
        LCB-1     & 0.0$\pm$0.0 \\
        EI-1      & 0.0$\pm$0.0 \\
        MACE-4    & 0.0$\pm$0.0 \\
        BLCB-1    & 0.0$\pm$0.0 \\
        EI-LP-4   & 0.0$\pm$0.0 \\
        \bottomrule
    \end{tabular}
\end{table}

@@

For this power amplifier, we aim to maximize the power added efficiency (PAE) and the output power (Pout), the $FOM$ is constructed as
$$
\mathit{FOM} = -3 \times \mathit{PAE} - \mathit{Pout}
$$

\begin{table}[htbp]
    \centering
    \caption{Optimization results of the power amplifier}
    \label{tab:result_PA}
    \begin{tabular}{llll}
        \toprule
        Algorithm & Results     \\ \midrule
        MACE-1    & 0.0$\pm$0.0 \\
        LCB-1     & 0.0$\pm$0.0 \\
        EI-1      & 0.0$\pm$0.0 \\
        MACE-4    & 0.0$\pm$0.0 \\
        BLCB-1    & 0.0$\pm$0.0 \\
        EI-LP-4   & 0.0$\pm$0.0 \\
        \bottomrule
    \end{tabular}
\end{table}

The MACE, BLCB, and EI-LP algorithms were tested in both sequential and batched
mode. The number of initial sampling is $N_{init} = 100$, the number of
iterations is $N_{iter} = 100$, the batch size is set to $B = 4$, so the total
number of HSPICE simulations is 500 for each batched run and 200 for each
sequential run.

@@

### 2018-02-05 paper/mace.tex
Paper update

Multi-objective optimization~\cite{MO_overview} can be formulated as:
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\bm{x}),~\dots~,f_m(\bm{x})
    \end{aligned}
\end{equation}

For multi-objective optimization, there are more than one objecive to optimize, and usually there does not exist a single solution that simultaneously optimize all these objectives. A design $\bm{x}_1$ is said to be \emph{dominate} by $\bm{x}_2$ if $\forall i \in \{1\dots m\},~f_i(\bm{x}_1) \le f_i(\bm{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bm{x}_1) < f_j(\bm{x}_2)$. A design is \emph{Pareto-optimal} if it is not dominated by any other point and dominate at least one point. The whole set of the non-dominated points in the design space is called the \emph{Pareto set}, and the set of non-dominated points in the objective space is called the \emph{Pareto front}. It is often unlikely to get the whole Pareto front as there might be infinite points on the front, the goal of multi-objective optimization is to find a set of designs that approximates the true Pareto front.

@@

We propose a novel heuristic for the parallelization of Bayesian optimization.
The parallelization is realized via multi-objective ensemble of multiple
acquisition functions. When we have multiple acquisition functions built from
the same GP model, they may not disagree with each other about which point is
most promising. For example, the value of LCB function always decreases as the
$\sigma(\bm{x})$ increases, however, for the PI function, when $\sigma(\bm{x})$
increases, the value of PI would decrease when $\mu(\bm{x}) < \tau$, and
increase when $\mu(\bm{x}) > \tau$.

% TODO: discuss the difference of acquisition functions: LCB: do not avoid repeated sampling, EI: always positive

With multi-objective optimization, the disagreement between different
acquisition functions can be used for the parallelization of Bayesian
optimization, as best trade-off between these acquisition functions can be
captured by the Pareto front.

In the proposed MACE algorithm, the LCB, EI, and PI acquisition functions are selected, but other acquisition functions like KG and PES can also be incorporated into the MACE framework. In each iteration, the following MO problem is constructed:
\begin{equation}
    \label{eq:MO_LCB_EI_PI}
    \begin{aligned}
        & \text{minimize} & & \mathrm{LCB}(\bm{x}),~-\mathrm{EI}(\bm{x}),~-\mathrm{PI}(\bm{x}).
    \end{aligned}
\end{equation}
Then the DEMO multi-objective optimization algorithm is applied to solve \eqref{eq:MO_LCB_EI_PI}. Once the Pareto front of LCB, EI and PI is obtained, the candidate evalution points are then randomly sampled from the Pareto front.

@@

\begin{algorithm}
\caption{Multi-objective Acquisition Ensemble Algorithm}
\label{alg:MACE}
\begin{algorithmic}[1]
\STATE Initial Sampling
\STATE Construct initial GP model
\FOR{t = 1, 2, \dots}
    \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
    \STATE Find the Pareto front of LCB, EI, PI function using the DEMO algorithm
    \STATE Randomly sample B points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto front where B is the batch size
    \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
    \STATE Update the GP model
\ENDFOR
\STATE Return best $f(\bm{x})$ recorded during iterations
\end{algorithmic}
\end{algorithm}

>>>>

Unlike single-objective optimization, multi-objective optimization problems
simultaneously optimize multiple objectives~\cite{MO_overview}, the
multi-objective optimization problem is formulated as
\begin{equation}
    \label{eq:MOFormulation}
    \begin{aligned}
        & \text{minimize} & & f_1(\bm{x}),~\dots~,f_m(\bm{x})
    \end{aligned}
\end{equation}
When we have multiple objectives to optimizes, the objectives can be conflicting, so that it is usually impossible to find a single solution that is the optimum of all objectives, the goal of multi-objective optimization algorithms is to appropriate the \emph{Pareto front} of the objectives. A solution $\bm{x}_1$ is said to \emph{dominate} $\bm{x}_2$ if $\forall i \in \{1\dots m\},~f_i(\bm{x}_1) \le f_i(\bm{x}_2)$ and $\exists j \in \{1\dots m\}, f_j(\bm{x}_1) < f_j(\bm{x}_2)$. A design is \emph{Pareto-optimal} if it is not dominated by any other point in the design space and dominate at least one point. The whole set of the Pareto-optimal points in the design space is called the \emph{Pareto set}, and the set of Pareto-optimal points in the objective space is called the \emph{Pareto front}. It is often unlikely to get the whole Pareto front as there might be infinite points on the front, the goal of multi-objective optimization is to find a set of designs that approximates the true Pareto front.

@@

We propose a novel heuristic for the parallelization of Bayesian optimization.
The parallelization is realized via the multi-objective ensemble of multiple
acquisition functions. As each acquisition function represents a unique
interpretation to the optimization problem, different acquisition functions may
not agree with each other about where to sample the next point. For example,
the value of LCB function always decreases as the $\sigma(\bm{x})$ increases,
however, for the PI function, when $\sigma(\bm{x})$ increases, the value of PI
would decrease when $\mu(\bm{x}) < \tau$, and increase when $\mu(\bm{x}) >
\tau$. For a noise-free problem, the EI values at already sampled points are
always zero, so that any unsampled point would have a better EI value than the
sampled points, while this property does not hold for the LCB function.

\textcolor{red}{TODO: Illustrate PF}
\begin{figure}[ht]
    \vskip 0.2in
    \begin{center}
        \centerline{\includegraphics[width=\columnwidth]{./img/pf.jpg}}
        \caption{Illustration of Pareto front}
        \label{fig:PF_example}
    \end{center}
    \vskip -0.2in
\end{figure}

With multi-objective optimization, the best trade-off between acquisition functions can be captured by the Pareto front of these acquisition functions. We can then sample on the Pareto front to obtain multiple candidate points for the objective function evaluations.

In the proposed MACE algorithm, the LCB, EI, and PI acquisition functions are selected, but other acquisition functions like KG and PES can also be incorporated into the MACE framework. In each iteration, the following MO problem is constructed:
\begin{equation}
    \label{eq:MO_LCB_EI_PI}
    \begin{aligned}
        & \text{minimize} & & \mathrm{LCB}(\bm{x}),~-\mathrm{EI}(\bm{x}),~-\mathrm{PI}(\bm{x}).
    \end{aligned}
\end{equation}
Then the DEMO multi-objective optimization algorithm is applied to solve the multi-objective problem in \eqref{eq:MO_LCB_EI_PI}. Once the Pareto front of LCB, EI and PI is obtained, the candidate evalution points are then randomly sampled from the Pareto front.

@@

\textcolor{red}{TODO: Input, Output parameters, see how other papers format the algorithm description}
\begin{algorithm}
\caption{Multi-objective Acquisition Ensemble Algorithm}
\label{alg:MACE}
\begin{algorithmic}[1]
\STATE Initial Sampling
\STATE Construct initial GP model
\FOR{t = 1, 2, \dots}
    \STATE Construct the LCB, EI and PI functions according to \eqref{eq:LCB} and \eqref{eq:PI_EI}
    \STATE Find the Pareto front of LCB, EI, PI function using the DEMO algorithm
    \STATE Randomly sample B points $\bm{x}_1, \dots, \bm{x}_B$ from the Pareto front where B is the batch size
    \STATE Evaluate $\bm{x}_1, \dots, \bm{x}_B$ to get $y_1 = f(\bm{x}_1),~\dots~,y_B = f(\bm{x}_B)$
    \STATE Update the GP model
\ENDFOR
\STATE Return best $f(\bm{x})$ recorded during iterations
\end{algorithmic}
\end{algorithm}

\textcolor{red}{TODO: If there is space, discuss the approximation of EI in a new subsection "Implementation details"}
